---
title: "Code for English Origin"
output: github_document
editor_options: 
  chunk_output_type: inline
---


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r install pkg}
install.packages('ggpubr')
```



```{r Load Package}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rstatix))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(rlang))
suppressPackageStartupMessages(library(gridExtra))



`%notin%` <- Negate(`%in%`)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```



```{r read data}
raw_data <- readxl::read_excel("/Users/shuyitan/Desktop/Consulting/Consulting_Linguistics_English_Origin/Data/Rap_data_Final.xlsx", sheet = 'MUEV+MIEV', skip = 1)
```



```{r}
str(raw_data)

```

Purpose: to investigate the use of English-origin verb

Dependent variable: 
MUEV or MIEV  

`# of syllables before` and `# of syllables after` can be ignored, because these two are already represented in the column `Relative position in phrase`  
0 = beginning of the phrase and 1 = end of the phrase


Following sound 
- Stop   
- Nasal 
------------
- Fricative   
- Affricate 
------------
- Approximant   
- Lateral approximant  
------------
- Vowel 
------------
- Zero   
These categories might need to be grouped. But there is a way to grouped them logically in linguistics. If they behave similarly and present to distinctive “behaviours”, here’s how they could be grouped: stop + nasal, fricative + affricate, stop + nasal + fricative + affricate, approximant + lateral approximant. Vowel and Zero need to stay alone, they can’t be grouped.  



Interested question: 
- if there are more MUEV in the Dictionnaire du Chilleur compared to the Fonds de données linguistiques du Québec  
- It might be interesting to look at the band `Dead Obies` separately from (or in comparison with) the other bands. They use the MUEV much more I think and they use more English words in their lyrics – they are typically known and criticized for that. 




For the factor Following Sound and the factor Final Sound of Verb (below in the table), the levels might need to be grouped. But there is a way to grouped them logically in linguistics. If they behave similarly and present to distinctive “behaviours”, here’s how they could be grouped: stop + nasal, fricative + affricate, stop + nasal + fricative + affricate, approximant + lateral approximant. Vowel and Zero need to stay alone, they can’t be grouped. 




```{r}
n_distinct(data$`Artist/band...1`)


n_distinct(data$Title)


nrow(data)


table(data$`MUEV or MIEV`)
```



```{r Count of Band}
raw_data %>% 
  group_by(`Artist/band...1`, ') %>%  
  tally() %>% 
  arrange(desc(n))
```



```{r count verb}
raw_data %>%  
  group_by(Verb) %>%  
  tally() %>%  
  arrange(desc(n)) 


raw_data %>%  
  group_by(`Regular or irregular V in Eng`) %>%  
  tally() %>%  
  arrange(desc(n)) 
```


```{r}
aggregated_words <- raw_data %>%
  group_by(Verb) %>%
  tally() %>%
  arrange(desc(n))

# Create a separate label for words with the same count
# Aggregate them into a single string (e.g., "word1, word2, ...")
aggregated_words <- aggregated_words %>%
  group_by(n) %>%
  summarise(words = paste(Verb, collapse = ", ")) %>%
  arrange(desc(n)) 



data_frame(aggregated_words)
aggregated_words

# aggregated_words <- aggregated_words %>%  
#   top_n(5)
# 
# # Plot the data
# ggplot(aggregated_words, aes(x = reorder(words, n), y = n)) +
#   geom_bar(stat = "identity") +
#   geom_text(aes(label = n), vjust = -0.5) + # Add count labels
#   coord_flip() + # Optional: to flip coordinates for better readability
#   labs(title = "Words with Top Counts",
#        x = "Words",
#        y = "Count") +
#   theme_minimal()
```



```{r relation position (density plot)}
rp_density <- ggplot(raw_data, aes(x = `Relative position in phrase`)) +
  geom_density(fill = "blue", alpha = 0.4) +
  labs(
    #title = "Density Plot of Relative Position in the Phrase",
    x = "Relative Position in Phrase",
    y = "Density"
  ) +
  theme_minimal() #+ 
  #theme(plot.subtitle = element_text(size = 5))

rp_density  
```



```{r}
# Convert the column to a factor with the desired order
raw_data <- raw_data %>%
  mutate(`Position of verb in phrase` = factor(
    `Position of verb in phrase`, 
    levels = c("beginning", "middle", "end")
  ))

# Calculate counts and percentages
data_summary <- raw_data %>%
  count(`Position of verb in phrase`) %>%
  mutate(percentage = round(n / sum(n) * 100)) 
```


```{r verb position (bar plot)}
# Create the bar plot
rp_bar <- ggplot(data_summary, aes(x = `Position of verb in phrase`, y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  geom_text(
    aes(label = paste0(n, " (", percentage, "%)")), 
    vjust = -0.5, 
    size = 4
  ) +
  labs(
    #title = "Bar Plot of Position of Verb in Phrase",
    x = "Position of Verb in Phrase",
    y = "Frequency"
  ) +
  theme_minimal() #+
  #theme(plot.subtitle = element_text(size = 5))

rp_bar
```



```{r following sound count}
raw_data %>% 
  group_by(`Following sound`) %>%  
  tally() 
```
check if significantly more 

grouping 

But there is a way to grouped them logically in linguistics. If they behave similarly and present to distinctive “behaviours”, here’s how they could be grouped: stop + nasal, fricative + affricate, stop + nasal + fricative + affricate, approximant + lateral approximant. Vowel and Zero need to stay alone, they can’t be grouped. 


```{r add following sound grouping}
data <- raw_data %>%  
  mutate(
    SoundGroup_1 = case_when(
      `Following sound` %in% c("stop", "nasal") ~ "stop + nasal",
      `Following sound` %in% c("fricative", "affricate") ~ "fricative + affricate",
      #`Following sound` %in% c("stop", "nasal", "fricative", "affricate") ~ "stop + nasal + fricative + affricate",
      `Following sound` %in% c("approximant", "lateral approximant") ~ "approximant + lateral approximant",
      `Following sound` %in% c("vowel", "zero") ~ `Following sound`, # Keep "vowel" and "zero" as is
      TRUE ~ "Other" # For any unexpected value
    ) ,
    SoundGroup_2 = case_when( 
      `Following sound` %in% c("stop", "nasal", "fricative", "affricate") ~ "stop + nasal + fricative + affricate", 
      `Following sound` %in% c("approximant", "lateral approximant") ~ "approximant + lateral approximant",
      `Following sound` %in% c("vowel", "zero") ~ `Following sound`, #  
      TRUE ~ "Other" # For any unexpected value
    )
  )

```

```{r additional followinh sound count}
data %>% 
  group_by(`SoundGroup_1`) %>%  
  tally()


data %>% 
  group_by(`SoundGroup_2`) %>%  
  tally()
```



```{r}



ggplot(data, aes(x = SoundGroup_1)) +
  geom_bar(fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Bar Chart of SoundGroup_1",
    x = "Sound Group",
    y = "Count"
  ) +
  theme_minimal() + 
    theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )



ggplot(data, aes(x = SoundGroup_2)) +
  geom_bar(fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Bar Chart of SoundGroup_1",
    x = "Sound Group",
    y = "Count"
  ) +
  theme_minimal() + 
    theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r}


# Precompute counts and percentages, and order by count
data_summary <- data %>%
  count(`Following sound`, name = "n") %>%
  mutate(
    percentage = n / sum(n) * 100, # Calculate percentage
    `Following sound` = factor(`Following sound`, levels = `Following sound`[order(-n)]) # Order factor
  )

# Create the bar plot with labels
ggplot(data_summary, aes(x = `Following sound`, y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black", alpha = 0.7) +
  geom_text(
    aes(label = paste0(n, " (", round(percentage, 1), "%)")), # Add count and percentage
    vjust = -0.5, size = 3
  ) +
  labs(
    title = "Following Sound",
    x = "Sound Group",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = -15, hjust = 0.2)
  )

```

```{r}
ggplot(data_summary, aes(x = SoundGroup_1, y = n, fill = `Following sound`)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8, color = "black") +
  geom_text(
    aes(
      label = ifelse(n > 5, paste0(`Following sound`, ": ", n), ""), # Only show labels if n > 5
      vjust = ifelse(n > 20, 0.5, -0.5) # Adjust vertical alignment for small segments
    ),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(
    title = "Stacked Bar Chart of Sound Groups with Counts",
    x = "Sound Group",
    y = "Count",
    fill = "Following Sound"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  )
```

```{r}
ggplot(data_summary, aes(y = SoundGroup_1, x = n, fill = `Following sound`)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8, color = "black") +
  geom_text(
    aes(
      label = ifelse(n > 5, paste0(`Following sound`, ": ", n), "")
    ),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(
    title = "Horizontal Stacked Bar Chart of Sound Groups with Counts",
    x = "Count",
    y = "Sound Group",
    fill = "Following Sound"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top"
  )
```


```{r}
ggplot(data_summary, aes(y = SoundGroup_1, x = n, fill = `Following sound`)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8, color = "black") +
  geom_text(
    aes(
      x = n + 1, 
      label = case_when(
        SoundGroup_1 == "fricative + affricate" & `Following sound` == "affricate" ~ paste0("affricate: ", n),
        SoundGroup_1 == "approximant + lateral approximant" & `Following sound` == "approximant" ~ paste0("approximate: ", n),
        n > 5 ~ paste0(`Following sound`, ": ", n), # Generic labels for other groups
        TRUE ~ "" # Do not label very small values
      )
    ),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(
    title = "Horizontal Stacked Bar Chart of Sound Groups with Customized Labels",
    x = "Count",
    y = "Sound Group",
    fill = "Following Sound"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top"
  )
```



```{r}
ggplot(data_summary, aes(y = SoundGroup_1, x = n, fill = `Following sound`)) + 
  geom_bar(stat = "identity", position = "stack", alpha = 0.8, color = "black") + 
  geom_text( 
    aes( 
      label = case_when( 
        SoundGroup_1 == "fricative + affricate" & `Following sound` == "affricate" ~ paste0("affricate: ", n), 
        SoundGroup_1 == "approximant + lateral approximant" & `Following sound` == "approximant" ~ paste0("approximate: ", n), 
        n > 5 ~ paste0(`Following sound`, ": ", n), # Generic labels for other groups 
        TRUE ~ "" # Do not label very small values 
      ), 
      x = case_when(
        `Following sound` == "affricate" ~ n + 18, # Move only affricate labels
        `Following sound` == "approximant" ~ n + 25, # Move only approximate labels
        TRUE ~ n # Default no adjustment
      ) 
    ), 
    position = position_stack(vjust = 0.5), 
    size = 3 
  ) + 
  labs( 
   # title = "Count of Following Sounds", 
    x = "Count", 
    y = "Sound Group", 
    fill = "Type of Following Sound" 
  ) + 
  theme_minimal() + 
  theme( 
    legend.position = "top" ,
    legend.text = element_text(size = 7.5)
  )
```



```{r other english item}
data %>%  
  group_by(`Presence of other English items in phrase`) %>% 
  tally() %>%  
  mutate(perc = n/sum(n))
```
check the pattern along with other linguistic factors 

```{r eng morphology}
data %>%  
  group_by(`Would have ENG morphology`) %>% 
  tally() %>% 
  mutate(perc = n/sum(n))
```
```{r}
data <- data %>%  
    mutate(
    `In Fonds de données linguistiques du Québec (FDLQ)` = as.logical(`In Fonds de données linguistiques du Québec (FDLQ)`)
  )


data %>%  
  summarize(     
    DC_TRUE = sum(`In Dictionnaire du Chilleur (DC)`),     
    FDLQ_TRUE = sum(`In Fonds de données linguistiques du Québec (FDLQ)`),     
    Both_TRUE = sum(`In Dictionnaire du Chilleur (DC)` & `In Fonds de données linguistiques du Québec (FDLQ)`),     
    Total = nrow(data) # Use nrow(data) instead of summing counts
  ) %>% 
  mutate(
    Percent_DC_TRUE = DC_TRUE / Total * 100,
    Percent_FDLQ_TRUE = FDLQ_TRUE / Total * 100,
    Percent_Both_TRUE = Both_TRUE / Total * 100,
    DC_Pct = paste0(DC_TRUE, " (", round(Percent_DC_TRUE, 2), "%)"),
    FDLQ_Pct = paste0(FDLQ_TRUE, " (", round(Percent_FDLQ_TRUE, 2), "%)"),
    Both_Pct = paste0(Both_TRUE, " (", round(Percent_Both_TRUE, 2), "%)")
  )

```



```{r}
data <- data %>%
  mutate(verb_def = case_when(
    `Verb category` == 1 ~ "simple verb",
    `Verb category` == 4 ~ "closed vowel + semivowel",
    `Verb category` == 5 ~ "final syllabic [l]",
    `Verb category` == 6 ~ "homophony, zero derivation, truncated, verlanized, or borrowed as is",
    TRUE ~ NA_character_  # Default case
  ))

data %>%
  group_by(`Verb category`, verb_def) %>%
  tally()  %>% 
  mutate(perc = n/nrow(data))

                  

```




```{r}
table_regularity <- table(data$`MUEV or MIEV`, data$`Regular or irregular V in Eng`)
chisq.test(table_regularity)
```

```{r}
fisher.test(table_regularity)
```

































